---
title: "Dats6101project"
author: "Zhe Wang"
date: "7/24/2020"
output:
  html_document:
    code_folding: hide
    # number_sections: true
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    toc: yes
---
```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = "markup", message = F)
# knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```
```{r outlierKD2}
# Fix outliers
outlierKD2 <- function(df, var, rm=FALSE) { 
    #' Original outlierKD functino by By Klodian Dhana,
    #' https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/
    #' Modified to have third argument for removing outliers instead of interactive prompt, 
    #' and after removing outlier, original df will not be changed. The function returns the a df, 
    #' which can be saved as original df name if desired.
    #' Also added QQ-plot in the output.
    #' Check outliers, and option to remove them, save as a new dataframe. 
    #' @param df The dataframe.
    #' @param var The variable in the dataframe to be checked for outliers
    #' @param rm Boolean. Whether to remove outliers or not.
    #' @return The dataframe with outliers replaced by NA if rm==TRUE, or df if nothing changed
    #' @examples
    #' outlierKD2(mydf, height, FALSE)
    #' mydf = outlierKD2(mydf, height, TRUE)
    #' mydfnew = outlierKD2(mydf, height, TRUE)
    dt = df # duplicate the dataframe for potential alteration
    var_name <- eval(substitute(var),eval(dt))
    na1 <- sum(is.na(var_name))
    m1 <- mean(var_name, na.rm = T)
    par(mfrow=c(2, 2), oma=c(0,0,3,0))
    boxplot(var_name, main="With outliers")
    hist(var_name, main="With outliers", xlab=NA, ylab=NA)
    qqnorm(var_name, main = "With outliers")
    qqline(var_name)
    outlier <- boxplot.stats(var_name)$out
    mo <- mean(outlier)
    var_name <- ifelse(var_name %in% outlier, NA, var_name)
    boxplot(var_name, main="Without outliers")
    hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
    qqnorm(var_name, main = "Without outliers")
    qqline(var_name)
    title("Outlier Check", outer=TRUE)
    na2 <- sum(is.na(var_name))
    cat("Outliers identified:", na2 - na1, "\n")
    cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "\n")
    cat("Mean of the outliers:", round(mo, 2), "\n")
    m2 <- mean(var_name, na.rm = T)
    cat("Mean without removing outliers:", round(m1, 2), "\n")
    cat("Mean if we remove outliers:", round(m2, 2), "\n")
    
    # response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
    # if(response == "y" | response == "yes"){
    if(rm){
        dt[as.character(substitute(var))] <- invisible(var_name)
        #assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
        cat("Outliers successfully removed", "\n")
        return(invisible(dt))
    } else {
        cat("Nothing changed", "\n")
        return(invisible(df))
    }
}
```
```{r xkablesummary}
loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")


xkabledply = function(modelsmmrytable, title="Table", digits = 4, pos="left", bso="striped", wide=FALSE) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display model summary. 
  #' wrapper for the base::summary function on model objects
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param modelsmmrytable This can be a generic table, a model object such as lm(), or the summary of a model object summary(lm()) 
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return HTML table for display
  #' @examples
  #' library("xtable")
  #' library("kableExtra")
  #' xkabledply( df, title="Table testing", pos="left", bso="hover" )
  if (wide) { modelsmmrytable <- t(modelsmmrytable) }
  modelsmmrytable %>%
    xtable() %>% 
    kable(caption = title, digits = digits) %>%
    kable_styling(bootstrap_options = bso, full_width = FALSE, position = pos)
}

xkablesummary = function(df, title="Table: Statistics summary.", digits = 4, pos="left", bso="striped") { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param df The dataframe.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return The HTML summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters, title="Five number summary", pos="left", bso="hover"  )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  xkabledply(s, title=title, digits = digits, pos=pos, bso=bso )
}

xkablevif = function(model, title="VIFs of the model", digits = 3, pos="left", bso="striped", wide=FALSE) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param model The lm or compatible model object.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return The HTML summary table of the VIFs for a model for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( lm(Salary~Hits+RBI, data=ISLR::Hitters, wide=T ) )
  
  vifs = table( names(model$coefficients)[2:length(model$coefficients)] ) # remove intercept to set column names
  vifs[] = faraway::vif(model) # set the values
  if (wide) { vifs <- t(vifs) }
  xkabledply( vifs, title=title, digits = digits, pos=pos, bso=bso )
}
```

##First, we import the data set wine.csv as wine, and check the structure. And then make are stored as factors, instead of numeric or integers.
```{r}
wine<-data.frame(read.csv("wine.csv"))
str(wine)
length(wine$quality)
wine$quality=as.factor(wine$quality)
str(wine)
xkablesummary(wine)
```


#Then, we want to remove the outliers. Firstly, we want to do the boxplot to see which variables have outliers, then we use outlierKD2 to remove the outliers.
```{r}
qqnorm(wine$fixed.acidity, main="qq plot of wine's fixed acidity")
qqnorm(wine$volatile.acidity,main="qqplot of wine's volatile.acidity")
qqnorm(wine$citric.acid,main="qqplot of wine's citric.acid")
qqnorm(wine$residual.sugar,main="qqplot of wine's residual.sugar")
qqnorm(wine$chlorides,main="qqplot of wine's chlorides")
qqnorm(wine$free.sulfur.dioxide,main="qqplot of wine's free.sulfur.dioxide")
qqnorm(wine$total.sulfur.dioxide,main="qqplot of wine's total.sulfur.dioxide")
qqnorm(wine$density,main="qqplot of wine's density")
qqnorm(wine$pH,main="qqplot of wine's pH",xlab="pH")
qqnorm(wine$sulphates,main="qqplot of wine's sulphates")
qqnorm(wine$alcohol,main="qqplot of wine's alcohol")
```

##After the qqplot, we could see that all these variables have outliers, so we use outlierKD2 function to remove the outliers. 
```{r}
wine<-outlierKD2(wine,fixed.acidity,TRUE)
wine<-outlierKD2(wine,volatile.acidity,TRUE)
wine<-outlierKD2(wine,citric.acid,TRUE)
wine<-outlierKD2(wine,residual.sugar,TRUE)
wine<-outlierKD2(wine,chlorides,TRUE)
wine<-outlierKD2(wine,free.sulfur.dioxide,TRUE)
wine<-outlierKD2(wine,total.sulfur.dioxide,TRUE)
wine<-outlierKD2(wine,density,TRUE)
wine<-outlierKD2(wine,pH,TRUE)
wine<-outlierKD2(wine,sulphates,TRUE)
wine<-outlierKD2(wine,alcohol,TRUE)
```

#Let's first do scatterplot matrix to select variables to included in our model
```{r}
names(wine)
pairs(~.,wine)
```
